Effectful Semantics for Substructural SSA

### 2023-11-05 pldi24/paper.tex
Tap
Literally 100 TODOs rip

\begin{lemma}[Semantic Upgrade]
  For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
  \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\hasty{\Gamma}{1}{a}{A}{q}}{}\]
\end{lemma}

@@

\TODO{State transformer on traces}

\TODO{Monoidal functor, yielding sliding}

@@

>>>>

\begin{lemma}[Semantic Upgrade]
  For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
  \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
\end{lemma}

@@

@@

Building on the ideas in \cite{mellies-ftrs}, we will show how a simple separation logic based on \cite{reynolds-separation-2002} slightly adapted to the categorical setting gives us a very rich \isotopessa model supporting a basic notion of refinement types, while at the same time giving us a nontrivial notion of nonlinear types.

We'll start by considering the category \(\ms{HRel}\), which we will take to have morphisms of the form \(A \to \ms{Heap} \to \mc{P}^+(A \times \ms{Heap} \times \ms{String}^* + \ms{String}^\omega)\), which is the result of applying the state-transformer monad with state \(\ms{Heap}\) to the trace monad with the infinite monoid of lists of strings described in the previous section.

\TODO{by the above, this is an \isotopessa model w/ nice properties; moving on!}

We now define a separation logic with predicates having the following grammar:
\begin{grammar}
  <\(\varphi, \psi\)> ::= 
    \(\varnothing\) 
    \;|\; \(e \mapsto a\) 
    \;|\; \(\varphi * \psi\)
    \;|\; \(\forall x: A, \varphi\)
    \;|\; \(\exists x: A, \varphi\)
    \;|\; \(\varphi \land \psi\)
    \;|\; \(\varphi \lor \psi\)
    \;|\; \(\varphi \Rightarrow \psi\)
    \;|\; \(\top\)
    \;|\; \(\bot\)
\end{grammar}
where \(e, a\) denote regular \isotopessa expressions and \(A, B\) regular \isotopessa types.

We can now do something somewhat tricky with \isotopessa dialects:
\TODO{truly pure expressions \(\implies\) semantics for expressions}
\TODO{giving us a semantics for separation logic stuff}
\TODO{now we define the category \(\ms{Sep}\) w/ objects contexts + predicates}
\TODO{tensor product, niceties \(\implies\) category, yey}
\TODO{sliding}
\TODO{therefore, substructural \textit{effectful} w/ nontrivial linear objects}
